RationalGRL: A Framework for Argumentation and Goal Modeling

### 2017-03-11 gmas.tex
merged thesis chapter with article

\section{Argument Schemes for Goal Modeling (GMAS)}
\label{sect:gmas}

In this section, we develop a set of argument schemes and critical questions for goal modeling. The final list of argument schemes and critical questions is shown in Table~\ref{table:argument-schemes}. The first four argument schemes (AS0-AS4) are argument for an element of a goal model, the next seven (AS5-AS11) are about relationships, the next two (AS12-AS13) are about intentional elements in general, and the last is (Att) is a generic counterargument for any type of argument that has been put forward. For each critical question, the right column shows the effect of answering the critical questions affirmatively, which can be \textsf{DISABLE} (disable the element of the original argument), \textsf{INTRO} (introduce a new element and argument), \textsf{REPLACE} (replace the element of the original argument), \textsf{ATTACK} (attack an argument directly). For instance, if the critical question CQ0 ``Is the actor relevant?'' is answered with ``No'', then the actor of the original argument AS0 will be disabled in the goal model.

%SG: If we only obtained them from these transcripts, how can we be sure that they are complete or even sufficient? Also, what are the transcripts? where did you take them from? why are they useful and why are they relevant?
We obtained the list in Table~\ref{table:argument-schemes} by analyzing transcripts of discussions between students about the development of an information system. In the first subsection, we provide details of this analysis process with concrete examples, after which we analyze our results in the next subsection. In the last subsection, we provide several examples of the effect of answering a critical question. In the next section, we develop a formal language for our argument schemes, critical questions, and the effect of answering them.

%SG: some comments on the table: softgoals can't decompose into goals or tasks? tasks don't contribute to goals? tasks mainly depend on resources.. why do we have A6 and A11? Note that GRL doesn't have those restrictions some languages such as i* have.. for example a softgoal can technically contribute to a task or a task can be decompsosed into goals.. how do you address those? We also mentioned in background that this is the benefit of GRL over i* so why do we now don't address it?

@@

@@

\paragraph{Experimental Setup} The assignment used for the experiments is to design a traffic simulator. Designers are provided a problem description, requirements, and a description of the desired outcomes. The original version of the problem descrption~\cite{UCIworkshop} is well known in the field of design reasoning since it has been used in a workshop\footnote{\url{http://www.ics.uci.edu/design-workshop/}}, and transcripts of this workshop have been analyzed in detail~\cite{Petre:2013:SDA:2535028}. Although the concepts of traffic lights, lanes, and intersections are common and appear to be simple, building a traffic simulator to represent these relationships and events in real time turns out to be challenging. Participants were asked to use a think-aloud method during the design session. The assignment was slightly adjusted to include several viewpoints as end products in order to conform to the course material~\cite{Bass:2012:SAP:2392670}. The final problem descriptions can be found in Appendix A of Schriek's master thesis~\cite{masterthesis1}. All groups were instructed to apply the \emph{functional architecture method}, focusing on developing the \emph{context}, the \emph{functional}, and the \emph{informational} viewpoints of the traffic simulator software. The students had two hours for the tasks, and the transcripts document the entire discussion. The details of the transcripts are shown in table~\ref{table:transcripts:info}.

@@

%SG: here we say we started from pras and then added more questions.. in the first paragraph we say the questions are coming from the transcripts!This is better... and we should mention it above..
We started with an initial list of 10 argument schemes and 18 critical questions that we derived from PRAS. We annotated transcripts with the arguments and critical questions from this list. If we found arguments or critical questions that did not appear in the original list, we added them and counted them as well. Argument schemes that did not appear were removed from the list, but critical questions were not removed (see discussion in Section~\ref{sect:gmas:transcripts:analysis}). Most of the occurrences were not literally found back, but had to be inferred from the context. This can be seen in the various examples we will discuss.

\todo{Marc}{Floris}{I'd like to say something in the next paragraph about the difficulty of extracting arguments from text. Do you know any references?}
It is generally known in the argumentation literature that it can be very difficult to annotate arguments correctly. Arguments are often imprecise, lack conclusion, and may be supported by non verbal communication that is not captured in the transcripts. Still, since research on argument extraction in the requirement engineering domains is in its infancy, we believe that our evaluation is useful by itself. Furthermore, our annotation is openly available\footnote{\todo{Marc}{Marc}{provide url}}, we provide parts of our annotation in Appendix~\ref{sect:transcripts:excerpts}, and most of the examples from this article come from the transcripts. In this way, we aim to make our annotation process as transparent as possible.

@@

@@

We annotated 41 applications of critical questions. Many critical questions (about 55\%) involved clarifying the name of an element, or discussing about the relevance of it (CQ12, CQ13).

@@

For each transcript, we manually create a GRL model from the argument schemes and critical questions we found in them, in order to verify whether the arguments put forward by the participants were sufficiently informative. An example of such a model is shown in Figure~\ref{fig:transcripts:grl}. We added green and red dots to various elements and relationships in the Figure. A green dot indicates there is an underlying argument for the element that is accepted, while a red dot indicates a rejected underlying argument. Note that if the underlying argument is rejected, the corresponding GRL element has been disabled. Some elements do not have a corresponding green or red dot. In that case, we have inferred the elements from the discussion, but we could not explicitly find back arguments for it.

We found that answering a critical questions can have four different effects on the original argument: \todo{Sepideh}{all}{Examples here are from the meeting scheduler! and our figure is about traffic simulator}
\begin{itemize}
\item \emph{INTRO}: Introduce a new goal element or relationship with a corresponding argument. This operation does not attack the original argument of the critical question, but rather creates a new argument. For instance, suppose argument scheme AS5 is instantiated as follows:: ``Goal \emph{Agreeable meeting dates} OR-decomposes into tasks \emph{Find agreeable date using scheduler} and \emph{Find agreeable date by talking to initiator}'' (Figure~\ref{fig:SchedulerModel}). Suppose now the critical question CQ5b: ``Does the goal \emph{Agreeable meeting dates} decompose into other tasks?'' is answered with ``yes, namely \emph{Find agreeable date by contacting secretary}''. This results in a new instantiation of AS5, namely: ``Goal \emph{Agreeable meeting dates} OR-decomposes into tasks \emph{Find agreeable date using scheduler}, \emph{Find agreeable date by talking to initiator}, and \emph{Find agreeable date by contacting secretary}. As a result, the goal model will contain the corresponding task \emph{Find agreeable date by contacting secretary}, as well as an OR-decomposition relation from the goal \emph{Agreeable meeting dates} to that task.
\item \emph{DISABLE:} Disable the element or relationship of the argument scheme to which the critical questions pertains. This operation does not create a new argument, but only disables (i.e., attacks) the original one. For instance, suppose argument scheme AS0 is instantiated with: ``Actor \emph{meeting initiator} is relevant'' (Figure~\ref{fig:SchedulerModel}). This argument can be attack with critical question CQ0: ``Actor \emph{meeting initiator} is not relevant''. As a result, the argument for the actor is attack, and actor \emph{meeting initiator} is disabled in the goal model.
\item \emph{REPLACE:} Replace the element of the argument scheme with a new element. This operation both introduces a new argument and attacks the original one. For instance, suppose argument scheme AS2 is instantiated wit: ``Actor \emph{meeting initiator} can perform task \emph{schedule meeting}. This argument can be attacked with critical question CQ13: ``The task \emph{schedule meeting} is unclear, it should be \emph{schedule board meeting}''. This results in replacing the original argument with the new argument ``Actor \emph{meeting initiator} can perform task \emph{schedule board meeting}. In the goal model, the description of the task should change accordingly.
\item \emph{ATTACK:} Attack any argument with an argument that cannot be classified as a critical question. For instance, suppose argument scheme AS4 is instantiated as follows: ``Actor \emph{meeting initiator} has goal \emph{quick}. For instance, suppose argument scheme AS0 is instantiated with ``Actor \emph{meeting scheduler} is relevant''. Suppose this argument is attacked with critical question CQ0: ``Actor \emph{meeting initiator} is not relevant'', because the participants decide to fully automate the meeting scheduling process. However, this critical question may in turn be attacked again, if new evidence comes up. For instance, it may turn out that it is sometimes more desirable to schedule meeting by hand. In this case, the generic counter argument ``Actor \emph{meeting initiator} is relevant, because it is sometimes easier to schedule meetings manually'' attack the argument ``Actor \emph{meeting scheduler} is not relevant''. As a result, the original argument ``Actor \emph{meeting scheduler} is relevant'' is accepted again, and as a result is shown in the goal model.
\end{itemize}

@@

\paragraph{Analysis of the Argument Schemes}
Our initial list of argument schemes consists of AS1-AS4, AS5-AS9 (Table~\ref{table:argument-schemes}). Therefore, the difference between the initial list of argument schemes and those found back in the transcripts is quite small. We found it surprising that we were able to find back all the schemes in the transcript at least twice, even more since the topic of discussion was not goal models, but more generally the architecture of an information system. This gives us an indication that these argument schemes are able to capture arguments used in those type of discussions to some extent.

More generally, we observed that our initial list is rather limited, which is a consequence of the fact that it is derived from PRAS. Since PRAS only considers very specific types of relationships, we are not able to capture many other relationships existing in GRL. GRL has four types of intentional elements (softgoal, goal, task, resource) and four types of relationships (positive contribution, negative contribution\footnote{In fact, a contribution can be any integer in the domain [-100,100], but for the sake of simplicity we only consider two kinds of contributions here.}, dependency, decomposition), allowing theoretically $4^3=64$ different types of argument schemes, of which we currently only consider 11. Our analysis however shows that many of these schemes are not often used, and thus, gives us some confidence in the resulting list. \todo{Sepideh}{all}{how does the analysis show that? also, what if someone says that you mentioned above that the transcripts were not used for goal models and that is why you didn't have those schemes.. what if someone is doing arguments for goal models, how do you make sure that the list is good enough, especially that many of the relationships are not considered?)}

\paragraph{Analysis of the Critical Questions} The difference between the initial list of critical questions and those we found back in the transcripts is much larger than for the critical questions. %SG: I don't get this sentence. Also, the ones below. 
On the one hand, we found a few of the critical questions we initially proposed. However, this does not mean that they were not implicitly used in the minds of the participants. If a participants makes an argument for a contribution from a task to a softgoal, it may very well be that she was asking herself the question ``Does the task contribute to some other softgoal?''. However, many of these critical questions are not mentioned explicitly. If we assume this explanation is at least partially correct, then, this would mean that critical questions may still play a role when formalizing the discussions leading up to a goal model, and it would be limiting to leave them out of our framework. In the context of tool support, we believe that having these critical questions available may stimulate discussions.

\subsection{Examples}
\label{sect:gmas:examples}

@@

\subsubsection{Example 1: Disable task \emph{Traffic light}}

The transcript excerpt of this example is shown in Table~\ref{table:transcripts:traffic-light} in the Appendix and comes from transcript $t_1$. In this example, participants are summing up functionality of the traffic simulator, which are tasks that the student can perform in the simulator. All these task can be formalized and are instantiations of argument scheme AS2: ``Actor \emph{Student} has tasks $T$'', where $T\in\{$Save map, Open map, Add intersection, Remove intersection, Add road, Add traffic light$\}$''.

@@

\subsubsection{Example 2: Clarify task \emph{Road pattern}}

The transcript excerpt of the second example is shown in Table~\ref{table:transcript:task-clarification} in the Appendix and comes from transcript $t_3$. It consists of a number of clarification steps, resulting in the task \texttt{Choose a road pattern}.

@@

What is clearly shown in this example is that a clarifying argument attacks all arguments previously used to describe the element. For instance, the final argument on the bottom of Figure~\ref{fig:examples:clarification} attacks all previous arguments. If this was not the case, then it may occur that a previous argument is \emph{reinstatiated}, meaning that it becomes accepted again because the argument attacking it is itself attacked.

\subsubsection{Example 3: Decompose goal \emph{Simulate}}

The transcript excerpt of this example is shown in Table~\ref{table:transcript:decomposition} in the Appendix and comes from transcript $t_3$. It consists of a discussion about the type of decomposition relationship for the goal \texttt{Simulate}.

@@

\subsubsection{Example 4: Reinstate actor \emph{Development team}}

The transcript excerpt of this example is shown in Table~\ref{table:transcript:irrelevant-actor} in the Appendix and comes from transcript $t_3$. It consists of two parts: first participant \texttt{P1} puts forth the suggestion to include actor \texttt{Development team} in the model. This is, then, questioned by participant \texttt{P2}, who argues that the professor will develop the software, so there won't be any development team. However, in the second part, participant \texttt{P2} argue that the development team should be considered, since the professor does not develop the software.

@@

We formalize this using a \emph{generic counterargument}, attacking the critical question. The first part of the discussion is shown in Figure~\ref{fig:examples:relevant-actor}. We formalize the first statement as an instantiation of argument scheme AS0: \texttt{Actor development team is relevant}. This argument is, then, attacked by answering critical question CQ0: \emph{Is actor development team relevant?} with \emph{No}. This results in two arguments, AS0 and CQ0, where CQ0 attacks AS0. This is shown in Figure~\ref{fig:examples:relevant-actor}, left image.

@@

Figure~\ref{fig:examples:relevant-actor2} shows the situation after the counter argument has been put forward. The argument ``The professor doesn't develop the software'' now attacks the argument ``\emph{Development team} is not relevant (\emph{The professor makes the software})'', which in turn attacks the original argument ``\emph{Development team} is relevant''. As a result, the first and the last argument are both accepted, which causes the actor in the GRL model to be enabled again.

>>>>

\section{Argument Schemes for Goal Modeling}
\label{sect:gmas}

@@

In this section, we develop a set of \emph{argument schemes for goal modeling} and associated critical questions. We start from an initial list that we derive from PRAS, containing argument schemes and critical questions that specific to elements and relationships of GRL. We then refine this list incrementally by annotating transcripts, meaning we identify and count instantiations of arguments schemes. If we find new argument schemes or critical questions, we add them to the list. If some of them do not occur at all, we remove them from the list. We then manually construct goal models from the arguments and counter arguments we find in the transcripts, and add traceability links between the goal model and underlying arguments. An example is shown in Figure~\ref{fig:transcripts:grl}, which is the same as the example in Figure~\ref{fig:trafficsim}, but now including tracability links to underlying arguments. A red dot indicates the underlying argument is rejected, and a green dot means the underlying argument is accepted.

The final list of argument schemes and critical questions that we end up with after our analysis is shown in Table~\ref{table:argument-schemes}. The initial list of argument schemes, i.e., those we start with before annotating the transcripts, consists of AS1-AS4, AS6-AS9 (Table~\ref{table:argument-schemes}), and their corresponding critical questions. The first four argument schemes (AS0-AS4) are arguments for an element of a goal model, the next seven (AS5-AS11) are about relationships, the next two (AS12-AS13) are about intentional elements in general, and the last is (Att) is a generic counterargument for any type of argument that has been put forward.

We found that answering critical questions can have varying effects on the model, and for each critical question, the right column in Table~\ref{table:argument-schemes} shows the effect of answering the critical questions affirmatively. Answering a critical questions can create an argument disabling the corresponding GRL element of the attacked argument scheme (\textsf{DISABLE}); it can create an argument introducing a new GRL element (\textsf{INTRO}); it can replace the GRL element corresponding to the original argument (\textsf{REPLACE}), or it can simply attack an argument directly (\textsf{ATTACK}).

In the first subsection of this section, we provide details of the transcript annotation process with concrete examples (see Appendix~\ref{sect:transcripts:excerpts} for transcript excerpts), after which we analyze our results in the second subsection.

@@

\paragraph{Experimental Setup} The assignment used for the experiments is to design a traffic simulator. Participants were asked to use a think-aloud method during the design session. The assignment was slightly adjusted to include several viewpoints as end products in order to conform to the course material~\cite{Bass:2012:SAP:2392670}. The full problem descriptions can be found in Appendix~\ref{ch:designprompt}. All groups were instructed to apply the \emph{functional architecture method}, focusing on developing the \emph{context}, the \emph{functional}, and the \emph{informational} viewpoints of the traffic simulator software. The students had two hours for the tasks, and the transcripts document the entire discussion. The details of the transcripts are shown in Table~\ref{table:transcripts:info}.

@@

We started with an initial list of 8 argument schemes and 18 critical questions that we derived from PRAS (AS1-AS4, AS6-AS9 of Table~\ref{table:argument-schemes}). We annotated transcripts with the arguments and critical questions from this list. If we found arguments or critical questions that did not appear in the original list, we added them and counted them as well. Argument schemes that did not appear were removed from the list, but critical questions were not removed (see discussion in Section~\ref{sect:gmas:transcripts:analysis}). Most of the occurrences were not literally found back, but had to be inferred from the context. This can be seen in the various examples we will discuss.

It is generally known in the argumentation literature that it can be very difficult to identify arguments in natural language texts~\cite{walton-etal2004}. Arguments are often imprecise, lack conclusion, and may be supported by non verbal communication that is not captured in the transcripts. However, there is hardly any research on on argument extraction in the requirement engineering domains, so despite this potential weakness in our approach, we believe it nevertheless is at least useful as something that others can build further on (see Section~\ref{sect:goalmodeling:openissues}).

@@

All original transcripts, annotations, and models are available on the following website:
\begin{quote} \url{http://www.github.com/marcvanzee/RationalArchitecture}
\end{quote}
in the folder ``Ch4 transcript analysis''. We also provide excerpts of the annotation in Appendix~\ref{sect:transcripts:excerpts}, and most of the examples we use in this article come from the transcripts.

@@

We annotated 41 applications of critical questions. Many critical questions (about 55\%) involved clarifying the name of an element, or discussing the relevance of it (CQ12, CQ13).

@@

For each transcript, we manually created a GRL model from the argument schemes and critical questions we found in them, in order to verify whether the arguments put forward by the participants were sufficiently informative. An example of such a model is shown in Figure~\ref{fig:transcripts:grl}. We added green and red dots to various elements and relationships in the figure. A green dot indicates there is an underlying argument for the element that is accepted, while a red dot indicates a rejected underlying argument. Note that if the underlying argument is rejected, the corresponding GRL element has been disabled. Some elements do not have a corresponding green or red dot. In that case, we have inferred the elements from the discussion, but we could not explicitly find back arguments for it.

We found that answering a critical questions can have four different effects on the original argument and the corresponding GRL element:
\begin{itemize}
\item \textsf{INTRO}: Introduce a new goal element or relationship with a corresponding argument. This operation does not attack the original argument of the critical question, but rather creates a new argument. For instance, suppose argument scheme AS5 is instantiated as follows:: ``Goal \texttt{Generate cars} OR-decomposes into tasks \texttt{Keep same cars} and \texttt{Create new cars}''. Suppose now the critical question CQ5b: ``Does the goal \texttt{Agreeable meeting dates} decompose into other tasks?'' is answered with ``yes, namely \texttt{Choose randomly}''. This results in a new instantiation of AS5, namely: ``Goal \texttt{Generate cars} OR-decomposes into tasks \texttt{Keep same cars}, \texttt{Create new cars}, and \texttt{Choose randomly}. As a result, the goal model will contain the corresponding task \texttt{Choose randomly}, as well as an OR-decomposition relation from the goal \texttt{Generate cars} to that task.
\item \textsf{DISABLE:} Disable the element or relationship of the argument scheme to which the critical questions pertains. This operation does not create a new argument, but only disables (i.e., attacks) the original one. For instance, suppose argument scheme AS0 is instantiated with: ``Actor \texttt{Teacher} is relevant''. This argument can be attack with critical question CQ0: ``Actor \texttt{Teacher} is not relevant''. As a result, the argument for the actor is attack, and actor \texttt{Teacher} is disabled in the goal model.
\item \textsf{REPLACE:} Replace the element of the argument scheme with a new element. This operation both introduces a new argument and attacks the original one. For instance, suppose argument scheme AS2 is instantiated with: ``Actor \texttt{Student} can perform task \texttt{Choose a pattern preference}. This argument can be attacked with critical question CQ13: ``The task \texttt{Choose a pattern preference} is unclear, it should be \texttt{Choose a road pattern}''. This results in replacing the original argument with the new argument ``Actor \texttt{Student} can perform task \texttt{Choose a road pattern}. In the goal model, the description of the task should change accordingly.
\item \textsf{ATTACK:} Attack any argument with an argument that cannot be classified as a critical question. For instance, suppose argument scheme AS0 is instantiated with ``Actor \texttt{Teacher} is relevant''. Suppose this argument is attacked with critical question CQ0: ``Actor \texttt{Teacher} is not relevant'', because she does not create the system. However, this critical question may in turn be attacked again, if new evidence comes up. For instance, it may turn out that the teacher will work on developing the application herself. In this case, the generic counter argument ``Actor \texttt{Teacher} is relevant, because does develop for the application'' attack the argument ``Actor \texttt{Teacher} is not relevant''. As a result, the original argument ``Actor \texttt{Teacher} is relevant'' is accepted again, and as a result is shown in the goal model.
\end{itemize}

@@

\paragraph{Analysis of the Argument Schemes}
Recall that our initial list of argument schemes consists of AS1-AS4, AS6-AS9 (Table~\ref{table:argument-schemes}). Therefore, the difference between the initial list of argument schemes and those found back in the transcripts is quite small. We found it surprising that we were able to find back all the schemes in the transcript at least twice, even more since the topic of discussion was not goal models, but more generally the architecture of an information system. This gives us an indication that it is possible to capture (parts of the) arguments used in those type of discussions using argument schemes.

We observed that our initial list is rather limited, which is a consequence of the fact that it is derived from PRAS. Since PRAS only considers very specific types of relationships, we are not able to capture many other relationships existing in GRL. GRL has four types of intentional elements (softgoal, goal, task, resource) and four types of relationships (positive contribution, negative contribution\footnote{In fact, a contribution can be any integer in the domain [-100,100], but for the sake of simplicity we only consider two kinds of contributions here.}, dependency, decomposition), allowing theoretically $4^3=64$ different types of argument schemes, of which we currently only consider 11. Our analysis however shows that many of these schemes are not often used, and thus, gives us some confidence in the resulting list. However, if needed, additional argument schemes and critical questions can be added, and our list is not meant to be exhaustive.

\paragraph{Analysis of the Critical Questions} The difference between the initial list of critical questions and those we found back in the transcripts is much larger than for the critical questions. %SG: I don't get this sentence. Also, the ones below. 
We found few of the critical questions we initially proposed. However, this does not mean that they were not implicitly used in the minds of the participants. If a participants for instance forms an argument for a contribution from a task to a softgoal, it may very well be that she was asking herself the question ``Does the task contribute to some other softgoal?''. However, many of these critical questions are not mentioned explicitly. If we assume this explanation is at least partially correct, then this would mean that critical questions may still play a role when formalizing the discussions leading up to a goal model, and it would be limiting to leave them out of our framework. In the context of tool support, we believe that having these critical questions available may stimulate discussions.

\section{Examples}
\label{sect:gmas:examples}

@@

\subsubsection{Example 1: Disable task \texttt{Traffic light}}

The transcript excerpt of this example is shown in Table~\ref{table:transcripts:traffic-light} in the appendix and comes from transcript $t_1$. In this example, participants are summing up functionality of the traffic simulator, which are tasks that the student can perform in the simulator. All these task can be formalized and are instantiations of argument scheme AS2: ``Actor \emph{Student} has tasks $T$'', where $T\in\{$Save map, Open map, Add intersection, Remove intersection, Add road, Add traffic light$\}$''.

@@

\subsubsection{Example 2: Clarify task \texttt{Road pattern}}

The transcript excerpt of the second example is shown in Table~\ref{table:transcript:task-clarification} in Appendix~\ref{sect:transcripts:excerpts} and comes from transcript $t_3$. It consists of a number of clarification steps, resulting in the task \texttt{Choose a road pattern}.

@@

What is clearly shown in this example is that a clarifying argument attacks all arguments previously used to describe the element. For instance, the final argument on the bottom of Figure~\ref{fig:examples:clarification} attacks all three other arguments for a name of the element. If this was not the case, then it may occur that a previous argument is \emph{reinstatiated}, meaning that it becomes accepted again because the argument attacking it is itself attacked. Suppose for instance the bottom argument ``Actor \texttt{Student} has task \texttt{Choose a pattern preference}'' did not attack the second argument: ``Action \texttt{Student} has task \texttt{Choose a pattern}''. In that case, this argument would be reinstated, because its only attacker ``Actor \texttt{Student} has task \texttt{Choose a pattern preference}'' is itself defeated by the bottom argument.

\subsubsection{Example 3: Decompose goal \texttt{Simulate}}

The transcript excerpt of this example is shown in Table~\ref{table:transcript:decomposition} in the appendix and comes from transcript $t_3$. It consists of a discussion about the type of decomposition relationship for the goal \texttt{Simulate}.

@@

\subsubsection{Example 4: Reinstate actor \texttt{Development team}}

The transcript excerpt of this example is shown in Table~\ref{table:transcript:irrelevant-actor} in the appendix and comes from transcript $t_3$. It consists of two parts: first participant \texttt{P1} puts forth the suggestion to include actor \texttt{Development team} in the model. This is, then, questioned by participant \texttt{P2}, who argues that the professor will develop the software, so there won't be any development team. However, in the second part, participant \texttt{P2} argue that the development team should be considered, since the professor does not develop the software.

@@

We formalize this using a \emph{generic counterargument}, attacking the critical question. The first part of the discussion is shown in Figure~\ref{fig:examples:relevant-actor}. We formalize the first statement as an instantiation of argument scheme AS0: ``Actor \texttt{development team} is relevant''. This argument is, then, attacked by answering critical question CQ0: ``Is actor \texttt{development team} relevant? with \emph{No}. This results in two arguments, AS0 and CQ0, where CQ0 attacks AS0. This is shown in Figure~\ref{fig:examples:relevant-actor}, left image.

@@

Figure~\ref{fig:examples:relevant-actor2} shows the situation after the counter argument has been put forward. The argument ``Actor \texttt{Professor} doesn't develop the software'' now attacks the argument ``\texttt{Development team} is not relevant (\emph{The professor makes the software})'', which in turn attacks the original argument ``\texttt{Development team} is relevant''. As a result, the first argument is reinstated, which causes the actor in the GRL model to be enabled again.

### 2017-03-11 main.tex
minor typo - test
formal framework
merged thesis chapter with article

\setcounter{tocdepth}{3}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{amssymb, amsmath}
\usepackage{array,url}
\usepackage{authblk}
\usepackage{caption}
\usepackage{courier}
\usepackage{enumitem}
\usepackage{graphicx}
\usepackage{helvet}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{sectsty}
\usepackage{subcaption}\captionsetup{compatibility=false}
\usepackage{tabularx}
\usepackage{tikz}\usetikzlibrary{arrows}
\usepackage{times}
\usepackage{url}
\usepackage{xcolor}

@@

@@

% use like this:
% \todo{floris}{marc}{Can you please update figure 2?} - Floris is asking Marc to update figure 2, this will show in red in the PDF
% \todo[resolved]{floris}{marc}{done} - Marc has updated fig 2 (or gives a reason why he doesn't want to uppdate the figure. This wiull show only in the tex

@@

\input{introduction}
\input{background}
\input{preview}
\input{gmas}
\input{formalframework}
\input{tool}
\input{methodology}
\input{relatedwork}
\input{conclusion}

@@

\input{transcripts}

>>>>

\setcounter{tocdepth}{3}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage[fleqn]{amsmath} % flush equations left
\usepackage{amssymb}
\usepackage{array,url}
\usepackage{authblk}
\usepackage{caption}
\usepackage{courier}
\usepackage{enumitem}
\usepackage{graphicx}
\usepackage{helvet}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{sectsty}
\usepackage{subcaption}\captionsetup{compatibility=false}
\usepackage{tabularx}
\usepackage{tikz}\usetikzlibrary{arrows, backgrounds}
\usepackage{times}
\usepackage{url}
\usepackage{xcolor}
\usepackage{fancyvrb}
\usepackage{algorithm}% http://ctan.org/pkg/algorithms
\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
\newcommand{\var}[1]{{\ttfamily#1}}% variable

@@

% Tikz stuff for arguments with GRL
\tikzset{
  argNodeIN/.style={shape=rectangle,draw,fill=white,align=center},
  argNodeOUT/.style={shape=rectangle, draw=black!60, text=black!40,fill=white},
  traceNodeRED/.style={circle,draw=black, fill=red, inner sep=0pt,minimum size=8pt},
  traceNodeGREEN/.style={circle,draw=black, fill=green, inner sep=0pt,minimum size=8pt},
  attackLink/.style={draw=black,->,line width=0.5mm},
  traceLinkRED/.style={draw, densely dotted,-,line width=0.5mm, draw=red},
  traceLinkGREEN/.style={draw, densely dotted,-,line width=0.5mm, draw=green},
  CQLink/.style={line width=0.5mm, dotted,->},
  ldots/.style={draw=none,rotate=90,font=\fontsize{20}{22.4}\selectfont},
  grl/.style={draw=none},
  disabled/.style={opacity=0.5}
}
\newcommand{\argtext}[2]{
\begin{tabularx}{0.2\textwidth}{X}
\ifx&#1&%
\else
\textbf{(#1)}
\fi
#2
\end{tabularx}
}
\newcommand{\smallargtext}[2]{
\begin{tabularx}{0.25\textwidth}{X}
\ifx&#1&%
\else
\textbf{(#1)}
\fi
#2
\end{tabularx}
}
\newcommand{\longargtext}[2]{
\begin{tabularx}{0.5\textwidth}{X}
\ifx&#1&%
\else
\textbf{(#1)}
\fi
#2
\end{tabularx}
}

% rationale
\newcommand{\rationale}{\emph{Rationale and example:}}

@@

% use like this:
% \todo{floris}{marc}{Can you please update figure 2?} - Floris is asking Marc to update figure 2, this will show in red in the PDF
% \todo[resolved]{floris}{marc}{done} - Marc has updated fig 2 (or gives a reason why he doesn't want to uppdate the figure. This will show only in the tex

@@

\input{introduction}
\input{background}
\input{overview}
\input{gmas}
\input{formalframework}
\input{tool}
\input{discussion}

@@

\input{transcripts}
\input{grl_specification}
